Component-Based mruby Platform for IoT Devices

### 2018-05-27 ipsj/chosha.tex
Final submit

\begin{biography}
\profile{Takuro Yamamoto}
{
received his B.E. degree from Osaka University in 2016.
His research interests include real-time and embedded systems\@.
}
\profile{Takuma Hara}
{
received his M.S. degree in Information Science from Nagoya University in 2013.
His research interests include embedded component technology\@.
}
\profile{Takuya Ishikawa}
{
received his Ph.D. degree in Information Science from Nagoya University in 2014.
His research interests include realtime scheduling theory, realtime operating system, and embedded component technology\@.
}
\profile{Hiroshi Oyama}
{
received Dr. Engineering degree from Gifu University in 2002.
He works at Okuma Corporation as Senior Engineer and also works at Nagoya University as Visiting Professor.
He is studying software components and programming language for embedded systems\@.
}
\profile{Hiroaki Takada}
{
is a professor at Institutes of Innovation for Future Society, Nagoya University.
He is also a professor and the Executive Director of the Center for Embedded Computing Systems (NCES), the Graduate School of Informatics, Nagoya University.
He received his Ph.D. degree in Information Science from University of Tokyo in 1996.
He was a Research Associate at University of Tokyo from 1989 to 1997, and was a Lecturer and then an Associate Professor at Toyohashi University of Technology from 1998 to 2003.
His research interests include real-time operating systems, real-time scheduling theory, and embedded system design.
He is a member of ACM, IEEE, IEICE, JSSST, and JSAE\@.
}
\profile{Takuya Azumi}
{
is an Associate Professor at the Graduate School of Science and Engineering, Saitama University.
He received his Ph.D. degree from the Graduate School of Information Science, Nagoya University. 
From 2008 to 2010, he was under the research fellowship for young scientists for Japan Society for the Promotion of Science. 
From 2010 to 2014, he was an Assistant Professor at the College of Information Science and Engineering, Ritsumeikan University. 
From 2014 to 2018, he was an Assistant Professor at the Graduate School of Engineering Science, Osaka University. 
His research interests include real-time operating systems and component-based development.
He is a member of IEEE, ACM, IPSJ, and IEICE\@.
}

>>>>

\begin{biography}
\profile{Takuro Yamamoto}
{
received his B.E. degree from Osaka University in 2016.
His research interests include real-time and embedded systems\@.
}
\profile{Takuma Hara}
{
received his M.S. degree in Information Science from Nagoya University in 2013.
His research interests include embedded component technology\@.
}
\profile{Takuya Ishikawa}
{
received his Ph.D. degree in information science from Nagoya University in 2013.
He was a researcher and then a designated assistant professor at the Center for Embedded Computing Systems, Nagoya University from 2014 to 2016.
His research interests include real-time scheduling theory, real-time operating systems, and embedded software platform\@.
}

\profile{Hiroshi Oyama}
{
received his Dr. Engineering degree from Gifu University in 2002.
He works at Okuma Corporation as Senior Engineer and also works at Nagoya University as Visiting Professor.
He is studying software components and programming language for embedded systems\@.
}
\profile{Hiroaki Takada}
{
is a professor at Institutes of Innovation for Future Society, Nagoya University.
He is also a professor and the Executive Director of the Center for Embedded Computing Systems (NCES), the Graduate School of Informatics, Nagoya University.
He received his Ph.D. degree in Information Science from University of Tokyo in 1996.
He was a Research Associate at University of Tokyo from 1989 to 1997, and was a Lecturer and then an Associate Professor at Toyohashi University of Technology from 1998 to 2003.
His research interests include real-time operating systems, real-time scheduling theory, and embedded system design.
He is a member of ACM, IEEE, IEICE, JSSST, and JSAE\@.
}
\profile{Takuya Azumi}
{
is an Associate Professor at the Graduate School of Science and Engineering, Saitama University.
He received his Ph.D. degree from the Graduate School of Information Science, Nagoya University. 
From 2008 to 2010, he was under the research fellowship for young scientists for Japan Society for the Promotion of Science. 
From 2010 to 2014, he was an Assistant Professor at the College of Information Science and Engineering, Ritsumeikan University. 
From 2014 to 2018, he was an Assistant Professor at the Graduate School of Engineering Science, Osaka University. 
His research interests include real-time operating systems and component-based development.
He is a member of IEEE, ACM, IPSJ, and IEICE\@.
}

### 2018-05-27 ipsj/paper.tex
Final submit

In embedded network software running on embedded systems within the Internet of Things (IoT), high levels of runtime efficiency and user productivity are required.
As an approach to improve the productivity of software development, the mruby on TOPPERS embedded component system (TECS) framework has been proposed; note that mruby on TECS framework employs a scripting language (i.e., a lightweight Ruby implementation) and supports component-based development.
In this paper, we propose an extended mruby on TECS framework for its application in developing software for IoT devices, including sensors and actuators.
Our proposed framework enables mruby programs to utilize Tomakomai Internetworking (TINET), a TCP/IP protocol stack specifically designed for use in embedded systems.
Further, the proposed framework incorporates two component-based functions, i.e., a componentized TINET stack called TINET+TECS and a componentized Two-Level Segregate Fit (TLSF) dynamic memory allocator called TLSF+TECS.
Here, TINET+TECS improves configurability and scalability and offers software developers high levels of productivity through variable network buffer sizes and the ability to add or remove various TCP (or UDP) functions.
TINET+TECS utilizes a dynamic TECS component connection method to satisfy the original TINET specifications.
Further, TLSF+TECS is a thread-safe memory allocator that runs at high speeds and efficiently consumes memory.
The experimental results of the comparison between TINET+TECS and the original TINET show that execution time and memory consumption overhead are both reduced; further, we conclude that configurability is improved.
Finally, the TLSF+TECS function which obtains and report statistical information regarding mruby's virtual machine (VM) memory usage, helps developers debug and verify their embedded IoT systems.

@@

    \item[R2:]
        Multiple mruby programs can run concurrently to improve productivity of software development.
        A thread-safe memory allocator is required to prevent the multiple mruby tasks from conflicting their memory.

@@

TECS statically performs component deployment and composition.
Consequently, connecting components does not incur significant overhead and memory requirements can be reduced.
TECS can be implemented in C, and demonstrates various feature such as source level portability and fine-grained components.

@@

mruby on TECS is a component-based framework for running an mruby script language on embedded systems.
This framework integrates two technologies, mruby and TECS, and enables to develop embedded software using a script language without slowing down the execution time.

\subsubsection{System Model of mruby on TECS}
% The present mruby on TECS system model is shown in Fig.\ref{fig:mrubyonTECS}.
Each mruby program, which is a bytecode, runs on own RiteVM as a componentized task of an RTOS.
TECS components support various embedded drivers such as motor and sensor drivers.
An mruby-TECS bridge provides native libraries for mruby and can call a native program (e.g., C legacy code) from an mruby program.
The mruby-TECS bridge also provides TECS components for receiving the invocation from an mruby program.

@@

Fig.\ref{fig:mrubyTECSbridge} illustrates an mruby-TECS bridge used to control a motor.
The left side of BridgeMotor belongs to the mruby program.
The right side of BridgeMotor belongs to TECS component.

@@

The mruby-TECS bridge generates a {\it celltype}, which is called from the mruby code, and an mruby class, which corresponds to a developer-specified TECS component to invoke a C function from the mruby program.
The generated mruby-TECS bridge supports registration of classes and methods for mruby.
Methods in an mruby class are defined by generation codes for an mruby-TECS bridge, such as setPower and stop.
Thus, when a method is called in an mruby program, the mruby-TECS bridge calls the function defined in the TECS component such as a Motor {\it cell}.

@@

The proposed framework is an extended mruby on TECS framework to develop network software for IoT devices.
Developers can use TCP- and UDP-related functions from mruby programs.
The proposed framework incorporates two functionalities: TINET+TECS and TLSF+TECS.
TINET+TECS is a component-based TCP/IP protocol stack comprised in the proposed framework, and it compensates for the original TINET's weak point that it is hard to maintain, extend, and analyze the software due to many complex source codes and improves the configurability.
TLSF+TECS which is a component-based dynamic memory allocator is used for memory management of RiteVMs and TCP/IP buffers in the proposed framework.
Since each TLSF component maintains its own heap area, TLSF+TECS allows concurrent operation without exclusive control while improving the efficiency of memory consumption, which is the advantage of TLSF.

@@

TINET is a compact TCP/IP protocol stack for embedded systems based on the ITRON\footnote{ITRON is an RTOS developed by the TRON project.} TCP/IP API Specification \cite{url:ITRON_TCP/IP_API_Spec}, developed by the TOPPERS Project \cite{url:TOPPERS}.
TINET has been released as an open-source tool.
To satisfy restrictions for embedded systems in terms of, for example, memory capacity, size, and power consumption, TINET supports the functions such as minimum copy frequency, elimination of dynamic memory control, asynchronous interfacing, error detailing per API.

@@

The original TINET eliminates dynamic memory control to meet the severe memory restrictions of embedded systems.
A memory area for sending/receiving data in the protocol stack is allocated and released within a predetermined area.
The memory allocator component allows for elimination of dynamic memory control in TINET+TECS by providing a requested memory area from the statically allocated memory area.

@@

Next, Fig.\ref{src:TLSFBuild} shows the build description of the TLSF memory allocator component illustrated in Fig.\ref{fig:WithTECS}\footnote{Other call/entry ports, attributes, and valuables are actually described, but it is omitted here for the simplicity.}.
Here, two sets of task components and TLSF components are combined.
Further, each memory pool size can be configured as a variable (lines 5 and 11 in Fig.\ref{src:TLSFBuild}).
Fig.\ref{src:TLSFC} presents the code that actually calls functions of the TLSF memory allocator component.
The use part shows a function in which the RiteVM allocates memory within the mruby on TECS framework\cite{par:mrubyonTECS}\cite{par:mrubyonTECS3}, which we introduced in Section \ref{sec:mrubyonTECS} above.
Line 8 calls the {\it free()} function of the TLSF memory allocator component.
{\it cMalloc\_} represents the name of the call port on line 2 in Fig.\ref{src:TLSFBuild}.
Similarly, lines 13 and 17 call the memory allocation function.
Particularly, the heap area for the {\it TLSFMalloc\_001} component is used if the code from \ref{src:TLSFC} is executed in {\it Task\_001}; conversely, if that code is executed in {\it Task\_002}, the heap area of the {\it TLSFMalloc\_002} component is used.
Using this approach, in component-based development using TECS, it is possible to operate with the same code without modifying the underlying C code, although the resulting cells differ.

@@

Fig.\ref{fig:EvaluationOfTLSFStatistics} shows the memory usage of two RiteVMs, with presented data acquired from the statistical information available via the TLSF+TECS framework.
From the figure, we observe that when the RiteVM are first activated, a large amount of memory is allocated for each RiteVM and initialization are performed within one second.
Next, the applications runs for ~10 s and the RiteVMs subsequently terminate.
The reduction in memory usage that occurs at regular intervals is due to the GC function of the RiteVM.
Here, an application that allocates tens of KB of memory runs on RiteVM 1 and an application that allocates hundreds of KB of memory runs on RiteVM 2.
Fig.\ref{fig:EvaluationOfMemoryUsageWithExclusiveContorol} shows the memory usage when two RiteVMs are executed in one heap area with exclusive control.
In Fig.\ref{fig:EvaluationOfDynamicConnection}, two applications are finished at 10.837 s, and in Fig.\ref{fig:EvaluationOfMemoryUsageWithExclusiveContorol}, they are finished at 10.877 s, which demonstrates that the execution time of the applications has increased due to exclusive control.
In addition to the overhead of execution time, a RiteVM is not affected by other RiteVMs since TLSF+TECS keeps heap area independently.
GC on a RiteVM is performed periodically; however, there is a part that GC is not performed periodically in RiteVM 1 of Fig.\ref{fig:EvaluationOfExecutionTime}.
In Fig.\ref{fig:EvaluationOfMemoryUsageWithExclusiveContorol}, two RiteVMs share one heap area; therefore, RiteVM 2 is affected by RiteVM 1.
On the other hand, in Fig.\ref{fig:EvaluationOfTLSFStatistics}, RiteVM 2 is not affected by RiteVM 1.

@@

Cloud-based platform reduces development efforts by providing cloud-base APIs and high-level constructs (e.g., drag-and-drop) \cite{par:cloud-based}.
In addition, it offers the ease deployment and evolution because the application logic is centrally located in a cloud platform.
However, it is platform-dependent design, and restricts developers in terms of functionality such as in-network aggregation or direct node-to-node communication locally.
The cloud-based mruby framework, enzi Board \cite{url:enzi}, has been proposed.
enzi can be developed and simulated on the Web, and developer can download and run the program on the board.

To address the issues of development efforts and platform-dependent design, MDD has been proposed \cite{par:MDD}.
MDD provides the benefits of reusable, platform-independent, extensible design, however it needs long development time to build MDD systems.

@@

The proposed framework can configure the TCP/IP protocol stack with minimum set compared to the other platforms.
Therefore, the proposed framework can reduce memory consumption.
In addition, the extended mruby on TECS supports legacy code such as a motor driver.
mruby has the same syntax as Ruby which has advantages to web application development as it uses in Rails framework \cite{url:rubyonrails}.

@@

This paper presented an extended framework of mruby on TECS, including TINET+TECS and TLSF+TECS.
In the proposed framework, mruby programs can call TINET+TECS functions through the mruby-TECS bridge.
Development of software for IoT devices such as sensors and actuators will be more efficient due to the high productivity of mruby.

TINET+TECS is a componentized version of TINET, a compact TCP/IP protocol stack that uses TECS.
It improves on TINET configurability while suppressing the overhead of componentization.
Scalability is also improved because the component-based framework simplifies to add/remove and change protocols such as TCP/UDP, IPv4/IPv6, and Ethernet/PPP.
This paper also presented the dynamic connection, a new TECS functionality, to enable dynamic processing while reducing memory consumption.
TINET+TECS utilizes the dynamic connection to satisfy the TINET specification for supporting the static generation of CEPs and REPs.
TLSF+TECS is a component-based dynamic allocator.
Since the TLSF+TECS can hold own heap area, memory contention will not occur even if memory is simultaneously allocated or released from multiple threads.
TINET+TECS and TLSF+TECS can be applicable to various embedded systems because these frameworks are executed on TECS and not limited to mruby.

>>>>

In embedded network software running on embedded systems within the Internet of Things (IoT), high levels of runtime efficiency and user productivity are required.
As an approach to improve the productivity of software development, the mruby on TOPPERS embedded component system (TECS) framework has been proposed; note that mruby on TECS framework employs a scripting language (i.e., a lightweight Ruby implementation) and supports component-based development.
In this paper, we propose an extended mruby on TECS framework for its application in developing software for IoT devices, including sensors and actuators.
Our proposed framework enables mruby programs to utilize Tomakomai Internetworking (TINET), a TCP/IP protocol stack specifically designed for use in embedded systems.
Further, the proposed framework incorporates two component-based functions, i.e., a componentized TINET stack called TINET+TECS and a componentized Two-Level Segregate Fit (TLSF) dynamic memory allocator called TLSF+TECS.
Here, TINET+TECS improves configurability and scalability and offers software developers high levels of productivity through variable network buffer sizes and the ability to add or remove various TCP (or UDP) functions.
TINET+TECS utilizes a dynamic TECS component connection method to satisfy the original TINET specifications.
Further, TLSF+TECS is a thread-safe memory allocator that runs at high speeds and efficiently consumes memory.
The experimental results of the comparison between TINET+TECS and the original TINET show that execution time and memory consumption overhead are both reduced; further, we conclude that configurability is improved.
Finally, the TLSF+TECS function which obtains and reports statistical information regarding mruby's virtual machine (VM) memory usage, helps developers debug and verify their embedded IoT systems.

@@

    \item[R2:]
        Multiple mruby programs can run concurrently to improve the productivity of software development.
        A thread-safe memory allocator is required to prevent the multiple mruby tasks from conflicting their memory.

@@

TECS statically performs component deployment and composition.
Consequently, connecting components does not incur significant overheads and memory requirements can be reduced.
TECS can be implemented in C, and demonstrates various features such as source level portability and fine-grained components.

@@

mruby on TECS is a component-based framework for running an mruby script language on embedded systems.
This framework integrates two technologies, mruby and TECS, and enables the development of embedded software using a script language without slowing down the execution time.

\subsubsection{System Model of mruby on TECS}
% The present mruby on TECS system model is shown in Fig.\ref{fig:mrubyonTECS}.
Each mruby program, which is a bytecode, runs on its own RiteVM as a componentized task of an RTOS.
TECS components support various embedded drivers such as motor and sensor drivers.
An mruby-TECS bridge provides native libraries for mruby and can call a native program (e.g., C legacy code) from an mruby program.
The mruby-TECS bridge also provides TECS components for receiving the invocation from an mruby program.

@@

Fig.\ref{fig:mrubyTECSbridge} illustrates an mruby-TECS bridge used to control a motor.
The left side of BridgeMotor belongs to the mruby program.
The right side of BridgeMotor belongs to the TECS component.

@@

The mruby-TECS bridge generates a {\it celltype}, which is called from the mruby code, and an mruby class, which corresponds to a developer-specified TECS component to invoke a C function from the mruby program.
The generated mruby-TECS bridge supports the registration of classes and methods for mruby.
Methods in an mruby class are defined by generation codes for an mruby-TECS bridge, such as setPower and stop.
Thus, when a method is called in an mruby program, the mruby-TECS bridge calls the function defined in the TECS component such as a Motor {\it cell}.

@@

The proposed framework is an extended mruby on TECS framework to develop network software for IoT devices.
Developers can use TCP- and UDP-related functions from mruby programs.
The proposed framework incorporates two functionalities: TINET+TECS and TLSF+TECS.
TINET+TECS is a component-based TCP/IP protocol stack comprised in the proposed framework, and it compensates for the original TINET's weak point that it is hard to maintain, extend, and analyze the software due to many complex source codes and improves the configurability.
TLSF+TECS which is a component-based dynamic memory allocator is used for the memory management of RiteVMs and TCP/IP buffers in the proposed framework.
Since each TLSF component maintains its own heap area, TLSF+TECS allows concurrent operation without exclusive control while improving the efficiency of memory consumption, which is the advantage of TLSF.

@@

TINET is a compact TCP/IP protocol stack for embedded systems based on the ITRON\footnote{ITRON is an RTOS developed by the TRON project.} TCP/IP API Specification \cite{url:ITRON_TCP/IP_API_Spec}, developed by the TOPPERS Project \cite{url:TOPPERS}.
TINET has been released as an open-source tool.
To satisfy restrictions for embedded systems in terms of, for example, memory capacity, size, and power consumption, TINET supports functions such as minimum copy frequency, elimination of dynamic memory control, asynchronous interfacing, error detailing per API.

@@

The original TINET eliminates dynamic memory control to meet the severe memory restrictions of embedded systems.
A memory area for sending/receiving data in the protocol stack is allocated and released within a predetermined area.
The memory allocator component allows for the elimination of dynamic memory control in TINET+TECS by providing a requested memory area from the statically allocated memory area.

@@

Next, Fig.\ref{src:TLSFBuild} shows the build description of the TLSF memory allocator component illustrated in Fig.\ref{fig:WithTECS}\footnote{Other call/entry ports, attributes, and valuables are actually described, but it is omitted here for simplicity.}.
Here, two sets of task components and TLSF components are combined.
Further, each memory pool size can be configured as a variable (lines 5 and 11 in Fig.\ref{src:TLSFBuild}).
Fig.\ref{src:TLSFC} presents the code that actually calls functions of the TLSF memory allocator component.
The use part shows a function in which the RiteVM allocates memory within the mruby on TECS framework\cite{par:mrubyonTECS}\cite{par:mrubyonTECS3}, which we introduced in Section \ref{sec:mrubyonTECS} above.
Line 8 calls the {\it free()} function of the TLSF memory allocator component.
{\it cMalloc\_} represents the name of the call port on line 2 in Fig.\ref{src:TLSFBuild}.
Similarly, lines 13 and 17 call the memory allocation function.
Particularly, the heap area for the {\it TLSFMalloc\_001} component is used if the code from \ref{src:TLSFC} is executed in {\it Task\_001}; conversely, if that code is executed in {\it Task\_002}, the heap area of the {\it TLSFMalloc\_002} component is used.
Using this approach, in component-based development using TECS, it is possible to operate with the same code without modifying the underlying C code, although the resulting cells differ.

@@

Fig.\ref{fig:EvaluationOfTLSFStatistics} shows the memory usage of two RiteVMs, with presented data acquired from the statistical information available via the TLSF+TECS framework.
From the figure, we observe that when the RiteVM are first activated, a large amount of memory is allocated for each RiteVM and initialization is performed within one second.
Next, the applications runs for ~10 s and the RiteVMs subsequently terminate.
The reduction in memory usage that occurs at regular intervals is due to the GC function of the RiteVM.
Here, an application that allocates tens of KB of memory runs on RiteVM 1 and an application that allocates hundreds of KB of memory runs on RiteVM 2.
Fig.\ref{fig:EvaluationOfMemoryUsageWithExclusiveContorol} shows the memory usage when two RiteVMs are executed in one heap area with exclusive control.
In Fig.\ref{fig:EvaluationOfDynamicConnection}, two applications are finished at 10.837 s, and in Fig.\ref{fig:EvaluationOfMemoryUsageWithExclusiveContorol}, they are finished at 10.877 s, which demonstrates that the execution time of the applications has increased due to exclusive control.
In addition to the overhead of execution time, a RiteVM is not affected by other RiteVMs since TLSF+TECS keeps heap area independently.
GC on a RiteVM is performed periodically; however, there is a part that GC is not performed periodically in RiteVM 1 of Fig.\ref{fig:EvaluationOfExecutionTime}.
In Fig.\ref{fig:EvaluationOfMemoryUsageWithExclusiveContorol}, two RiteVMs share one heap area; therefore, RiteVM 2 is affected by RiteVM 1.
On the other hand, in Fig.\ref{fig:EvaluationOfTLSFStatistics}, RiteVM 2 is not affected by RiteVM 1.

@@

A cloud-based platform reduces development efforts by providing cloud-base APIs and high-level constructs (e.g., drag-and-drop) \cite{par:cloud-based}.
In addition, it offers easy deployment and evolution because the application logic is centrally located in a cloud platform.
However, it is a platform-dependent design, and restricts developers in terms of functionality such as in-network aggregation or direct node-to-node communication locally.
The cloud-based mruby framework, enzi Board \cite{url:enzi}, has been proposed.
enzi can be developed and simulated on the Web, and the developer can download and run the program on the board.

To address the issues of development efforts and platform-dependent design, MDD has been proposed \cite{par:MDD}.
MDD provides the benefits of reusable, platform-independent, extensible design, however it needs a long development time to build MDD systems.

@@

The proposed framework can configure the TCP/IP protocol stack with minimum set compared to the other platforms.
Therefore, the proposed framework can reduce memory consumption.
In addition, the extended mruby on TECS supports a legacy code such as a motor driver.
mruby has the same syntax as Ruby which has advantages for web application development as it uses in Rails framework \cite{url:rubyonrails}.

@@

This paper presented an extended framework of mruby on TECS, including TINET+TECS and TLSF+TECS.
In the proposed framework, mruby programs can call TINET+TECS functions through the mruby-TECS bridge.
The development of software for IoT devices such as sensors and actuators will be more efficient due to the high productivity of mruby.

TINET+TECS is a componentized version of TINET, a compact TCP/IP protocol stack that uses TECS.
It improves on TINET configurability while suppressing the overhead of componentization.
Scalability is also improved because the component-based framework simplifies to add/remove and change protocols such as TCP/UDP, IPv4/IPv6, and Ethernet/PPP.
This paper also presented the dynamic connection, a new TECS functionality, to enable dynamic processing while reducing memory consumption.
TINET+TECS utilizes the dynamic connection to satisfy the TINET specification for supporting the static generation of CEPs and REPs.
TLSF+TECS is a component-based dynamic allocator.
Since the TLSF+TECS can hold its own heap area, memory contention will not occur even if memory is simultaneously allocated or released from multiple threads.
TINET+TECS and TLSF+TECS can be applicable to various embedded systems because these frameworks are executed on TECS and not limited to mruby.
